Pool Model - A unified Mathematical\\ Modelling Framework for Food Science

### 2025-09-03 paper/pool_model.tex
remove listings from appendix

\begin{table}
    \centering
    \begin{tabularx}{\textwidth}{@{}l *2{>{\centering\arraybackslash}X} l *2{>{\centering\arraybackslash}X}@{}}
        \multicolumn{3}{c}{\textbf{\acs{ode}}} & \multicolumn{3}{c}{\textbf{\acs{abm}}}\\
        \toprule
        \multicolumn{6}{c}{\textbf{Cellular Reactions}}\\
        \toprule
        Production $G_A$            & $\alpha_A$    & $0.002$\unit{\per\minute}
        & Uptake                    & $u_A$         & $0.01$\unit{\per\minute}\\
        Production $G_B$            & $\alpha_B$    & $0.002$\unit{\per\minute}
        & Uptake                    & $u_B$         & $0.01$\unit{\per\minute}\\
        Production $I$              & $\kappa$      & $0.1$\unit{\per\minute}
        & Production $I$            & $b$           & $0.1$\unit{\per\minute}\\
        Inhibition                  & $\mu_I$       & $1.0\cdot 10^{-7}$\unit{\per\micro\metre\cubed}
        &                           & $\nu_I$       & $0.1$\unit{\micro\metre\cubed\per\mol}\\
        &&
        & Conversion                & $\sigma$      & $0.1$\unit{\micro\metre\cubed\per\mol}\\
        Total volume                & $N_t$& $2.0\cdot 10^5$\unit{\micro\metre\cubed}
        &&&\\
        &&
        & Diffusion $R$             & $D_R$         & $30.0$\unit{\micro\metre\squared\per\minute}\\
        &&
        & Diffusion $I$             & $D_I$         & $30.0$\unit{\micro\metre\squared\per\minute}\\
        % Note: this diffusion roughly corresponds to the one of ethanol in water
        % (see https://de.wikipedia.org/wiki/Diffusionskoeffizient#In_Fl%C3%BCssigkeiten)
        % 15.0 µm^2/min = 0.9 m^2/s
        % Ethanol: 0.84 m^2/s
        \toprule
        \multicolumn{6}{c}{\textbf{Physical Mechanics \& Interactions}}\\
        \toprule
        &&& Damping                 & $\eta$        & $0.5$\unit{\per\minute}\\
        &&& Mass                    & $m$           & $7.70$\unit{\pico\gram}\\
        &&& Potential               & $\epsilon$    & $0.5$\unit{\pico\gram\micro\metre\per\minute\squared}\\
        \toprule
        \multicolumn{6}{c}{\textbf{Cell Cycle}}\\
        \toprule
        Rate $L_A\rightarrow G_A$   & $\lambda_A$   & $0.0025$\unit{\per\minute}
        &                           & $\lambda_A$   & $0.0025$\unit{\per\minute}\\
        Rate $L_B\rightarrow G_B$   & $\lambda_B$   & $0.0005$\unit{\per\minute}
        &                           & $\lambda_B$   & $0.0005$\unit{\per\minute}\\
        &&& Division Vol.           & $V_t$         & $14.14$\unit{\micro\metre\cubed}\\
        \toprule
        \multicolumn{6}{c}{\textbf{Initial Values}}\\
        \toprule
        Volume of $L_A$             & $L_A$         & $127.2$\unit{\micro\metre\cubed}
        & Cell Number               & $C_A$         & 18\\
        Volume of $G_A$             & $G_A$         & $0.0$\unit{\micro\metre\cubed}
        & Cell Volume               & $V_{c,A}$     & $7.07$\unit{\micro\metre\cubed}\\
        Volume of $L_B$             & $L_B$         & $127.23$\unit{\micro\metre\cubed}
        & Cell Number               & $C_B$         & 18\\
        Volume of $G_B$             & $G_B$         & $0.0$\unit{\micro\metre\cubed}
        & Cell Volume               & $V_{c,B}$     & $7.07$\unit{\micro\metre\cubed}\\
        Relative Nutrients          & $R$         & $1.0$
        & Nutrients                 & $[R_0]$       & $1.0$\unit{\mol\per\micro\metre\cubed}\\
        Inhibitor                   & $I_0$         & $0.0$\unit{\mol}
        & Inhibitor                 & $[I_0]$       & $0.0$\unit{\mol\per\micro\metre\cubed}\\
        \bottomrule
    \end{tabularx}
    \caption{
        Comparison of the parameters for the \ac{ode} model and \ac{abm}.
        % To view, how these parameters are stored inside a cell-agent, we refer to the supplementary
        % material~\ref{sec:supplement-full-parameters}.
    }
    \label{tab:spatial_limit_init}
\end{table}

\begin{listing}
\begin{lstlisting}[language=json]
{
    "diffusion_constants": [
        15.0,
        15.0
    ],
    "initial_concentrations": [
        2.0,
        0.0
    ],
    "size": 1000.0,
    "n_voxels": null
}
\end{lstlisting}
\caption{
    The \texttt{domain.json} file specifying initial values of the domain.
}
\end{listing}
%
\begin{listing}
\begin{lstlisting}[language=json]
{
    "random_seed": 1,
    "n_times": 20001,
    "dt": 0.5,
    "t_start": 0.0,
    "save_interval": 500,
    "n_threads": 8,
    "show_progressbar": true,
    "save_path": "out/pool_model",
    "save_add_date": true
}
\end{lstlisting}
\caption{
    The \texttt{meta\_params.json} file specifying all parameters not directly related to the simulation content.
    These settings do not represent any model-specific properties but rather generic ones such as time-span or number of threads.
}
\end{listing}
%
\begin{listing}
\begin{lstlisting}[language=json]
{
    "mechanics": {
        "pos": [
        142.85714285714286,
        142.85714285714286
        ],
        "vel": [
        0.0,
        0.0
        ],
        "damping_constant": 0.5,
        "mass": 7.704755982928968
    },
    "cycle": {
        "volume_division_threshold": 14.137166941154069,
        "lag_phase_transition_rate_1": 0.005,
        "lag_phase_transition_rate_2": 0.0025
    },
    "cellular_reactions": {
        "potential_strength": 0.5,
        "food_to_volume_conversion": 0.1,
        "lag_phase_active": true,
        "species": "S1",
        "cell_volume": 7.0685834705770345,
        "uptake_rate": 0.01,
        "inhibition_production_rate": 0.1,
        "inhibition_coefficient": 0.1
    },
    "interactionextracellulargradient": null
}
\end{lstlisting}
\caption{
    The first entry of the \texttt{initial\_cells.json} file which defines every individual particle present in the simulation initially.
}
\end{listing}
%
%
% ................................................................................................
\subsubsection{Homogenous and Inhomogenous Initial Cell Distributions}
To clarify what we mean by homogenous and inhomogenous initial configurations, we include snapshots of the simulation.
At this point, cells have already grown such that the sizes of cell-clumps are better visible by the naked eye.
The snapshots were taken after $2.08$\unit{\day}.
\begin{figure}
    \centering
    \begin{subfigure}[c]{0.5\textwidth}
        \includegraphics[width=\textwidth]{Figures/abm-homogenous/snapshot_00012000.png}
        \subcaption{Homogenous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
    \end{subfigure}%
    \begin{subfigure}[c]{0.5\textwidth}
        \includegraphics[width=\textwidth]{Figures/abm-homogenous-low-diffusion/snapshot_00012000.png}
        \subcaption{Homogenous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
    \end{subfigure}\\
    \begin{subfigure}[c]{0.5\textwidth}
        \includegraphics[width=\textwidth]{Figures/abm-inhomogenous/snapshot_00012000.png}%
        \subcaption{Heterogeneous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
    \end{subfigure}%
    \begin{subfigure}[c]{0.5\textwidth}
        \includegraphics[width=\textwidth]{Figures/abm-random/snapshot_00012000.png}%
        \subcaption{Random initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
    \end{subfigure}
    \caption{
        Spatial configuration of cells corresponding to the pictures shown in figure~\ref{fig:spatial-growth-curve}.
    }
    \label{fig:supplement-spatial-initial-config-cells}
\end{figure}
%
%
\end{document}

>>>>

\begin{table}
    \begin{tabularx}{\textwidth}{@{}l *2{>{\centering\arraybackslash}X} l *2{>{\centering\arraybackslash}X}@{}}
        \multicolumn{3}{c}{\textbf{\acs{ode}}} & \multicolumn{3}{c}{\textbf{\acs{abm}}}\\
        \multicolumn{6}{c}{\textbf{Cellular Reactions}}\\
        Production $G_A$            & $\alpha_A$    & $0.002$\unit{\per\minute}
        & Uptake                    & $u_A$         & $0.01$\unit{\per\minute}\\
        Production $G_B$            & $\alpha_B$    & $0.002$\unit{\per\minute}
        & Uptake                    & $u_B$         & $0.01$\unit{\per\minute}\\
        Production $I$              & $\kappa$      & $0.1$\unit{\per\minute}
        & Production $I$            & $b$           & $0.1$\unit{\per\minute}\\
        Inhibition                  & $\mu_I$       & $1.0\cdot 10^{-7}$\unit{\per\micro\metre\cubed}
        &                           & $\nu_I$       & $0.1$\unit{\micro\metre\cubed\per\mol}\\
        &&
        & Conversion                & $\sigma$      & $0.1$\unit{\micro\metre\cubed\per\mol}\\
        Total volume                & $N_t$& $2.0\cdot 10^5$\unit{\micro\metre\cubed}
        &&&\\
        &&
        & Diffusion $R$             & $D_R$         & $30.0$\unit{\micro\metre\squared\per\minute}\\
        &&
        & Diffusion $I$             & $D_I$         & $30.0$\unit{\micro\metre\squared\per\minute}\\
        % Note: this diffusion roughly corresponds to the one of ethanol in water
        % (see https://de.wikipedia.org/wiki/Diffusionskoeffizient#In_Fl%C3%BCssigkeiten)
        % 15.0 µm^2/min = 0.9 m^2/s
        % Ethanol: 0.84 m^2/s
        \multicolumn{6}{c}{\textbf{Physical Mechanics \& Interactions}}\\
        &&& Damping                 & $\eta$        & $0.5$\unit{\per\minute}\\
        &&& Mass                    & $m$           & $7.70$\unit{\pico\gram}\\
        &&& Potential               & $\epsilon$    & $0.5$\unit{\pico\gram\micro\metre\per\minute\squared}\\
        \multicolumn{6}{c}{\textbf{Cell Cycle}}\\
        Rate $L_A\rightarrow G_A$   & $\lambda_A$   & $0.0025$\unit{\per\minute}
        &                           & $\lambda_A$   & $0.0025$\unit{\per\minute}\\
        Rate $L_B\rightarrow G_B$   & $\lambda_B$   & $0.0005$\unit{\per\minute}
        &                           & $\lambda_B$   & $0.0005$\unit{\per\minute}\\
        &&& Division Vol.           & $V_t$         & $14.14$\unit{\micro\metre\cubed}\\
        \multicolumn{6}{c}{\textbf{Initial Values}}\\
        Volume of $L_A$             & $L_A$         & $127.2$\unit{\micro\metre\cubed}
        & Cell Number               & $C_A$         & 18\\
        Volume of $G_A$             & $G_A$         & $0.0$\unit{\micro\metre\cubed}
        & Cell Volume               & $V_{c,A}$     & $7.07$\unit{\micro\metre\cubed}\\
        Volume of $L_B$             & $L_B$         & $127.23$\unit{\micro\metre\cubed}
        & Cell Number               & $C_B$         & 18\\
        Volume of $G_B$             & $G_B$         & $0.0$\unit{\micro\metre\cubed}
        & Cell Volume               & $V_{c,B}$     & $7.07$\unit{\micro\metre\cubed}\\
        Relative Nutrients          & $R$         & $1.0$
        & Nutrients                 & $[R_0]$       & $1.0$\unit{\mol\per\micro\metre\cubed}\\
        Inhibitor                   & $I_0$         & $0.0$\unit{\mol}
        & Inhibitor                 & $[I_0]$       & $0.0$\unit{\mol\per\micro\metre\cubed}\\
        Comparison of the parameters for the \ac{ode} model and \ac{abm}.
        % To view, how these parameters are stored inside a cell-agent, we refer to the supplementary
        % material~\ref{sec:supplement-full-parameters}.
    \label{tab:spatial_limit_init}
\end{table}
\subsubsection{Homogenous and Inhomogenous Initial Cell Distributions}
To clarify what we mean by homogenous and inhomogenous initial configurations, we include snapshots of the simulation.
At this point, cells have already grown such that the sizes of cell-clumps are better visible by the naked eye.
The snapshots were taken after $2.08$\unit{\day}.
    \begin{subfigure}[c]{0.5\textwidth}
        \includegraphics[width=\textwidth]{Figures/abm-homogenous/snapshot_00012000.png}
        \subcaption{Homogenous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
    \begin{subfigure}[c]{0.5\textwidth}
        \includegraphics[width=\textwidth]{Figures/abm-homogenous-low-diffusion/snapshot_00012000.png}
        \subcaption{Homogenous initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
    \begin{subfigure}[c]{0.5\textwidth}
        \includegraphics[width=\textwidth]{Figures/abm-inhomogenous/snapshot_00012000.png}%
        \subcaption{Heterogeneous initial cell distribution, $D_R=D_I=30.0$\unit{\micro\metre\squared\per\minute}}
    \begin{subfigure}[c]{0.5\textwidth}
        \includegraphics[width=\textwidth]{Figures/abm-random/snapshot_00012000.png}%
        \subcaption{Random initial cell distribution, $D_R=D_I=5.0$\unit{\micro\metre\squared\per\minute}}
        Spatial configuration of cells corresponding to the pictures shown in figure~\ref{fig:spatial-growth-curve}.
    \label{fig:supplement-spatial-initial-config-cells}
\end{document}
